Data Exploration: IUCN and AquaMaps

### 2015-09-04 data_explore.Rmd
added regression analysis

spp_list = spp_list%>%
              mutate(am_area = NA,
                     iucn_area=NA,
                     perc = NA,
                     am_perc=NA,
                     iucn_perc=NA,
                     sm_perc = NA, #filled with % of smaller range that is within the larger range
                     sm_range = NA) #what range is smaller

@@

  #get just the aquamaps cells
  am_spp_map = spp_map%>%
                  filter(!is.na(am_prob))%>%
                dplyr::select(loiczid,am_sid,am_prob)%>% #add in am_prob when we want to look at probability
                  unique()
  #rasterize aquamaps map
  r_am_spp  <-  subs(loiczid_raster, 
                     am_spp_map[ , c('loiczid', 'am_sid')], 
                     by = 'loiczid', 
                     which = 'am_sid', 
                     subsWithNA = TRUE)
#   
#   spp_list[i,]$am_area = area(r_am_spp,na.rm=T)%>%
#               cellStats(.,stat='sum')
#

  #get the iucn cells
  iucn_map <- spp_map %>%
    mutate(iucn_pres = iucn_area > 0)%>%
      dplyr::select(loiczid,iucn_area,am_sid,am_prob,iucn_pres)%>%
        unique()

@@

   if(TRUE %in% duplicated(iucn_map$loiczid))next()
#   
#  rasterize IUCN map
  r_iucn_spp <- subs(loiczid_raster, 
                     iucn_map[ , c('loiczid', 'iucn_pres')], 
                    by = 'loiczid', 
                     which = 'iucn_pres', 
                     subsWithNA = TRUE)
#   
#   spp_list[i,]$iucn_area = area(r_iucn_spp,na.rm=T)%>%
#                 cellStats(.,stat='sum')

@@

  #iucn map cells
  cells_iucn<-iucn_map%>%
              filter(iucn_pres==TRUE)%>%
              .$loiczid
  cells_am <- am_spp_map%>%
                filter(!is.na(am_sid))%>%
                .$loiczid

  cells_total = unique(spp_map$loiczid)
  cells_overlap = intersect(cells_iucn,cells_am)

  #percent overlap
  perc = (length(cells_overlap)/length(cells_total))*100
  spp_list[i,]$perc = perc

  am_only = setdiff(cells_am,cells_iucn)
  iucn_only = setdiff(cells_iucn,cells_am)

  perc_am = (length(am_only)/length(cells_total))*100
  perc_iucn = (length(iucn_only)/length(cells_total))*100

  spp_list[i,]$am_perc = perc_am
  spp_list[i,]$iucn_perc = perc_iucn

  spp_list[i,]$sm_range = ifelse(length(cells_am)<length(cells_iucn),'AM','IUCN') #what dataset has the smaller range
  spp_list[i,]$sm_perc = length(cells_overlap)/min(length(cells_am),length(cells_iucn))*100 #this is the total percent of smaller range encompassed in the larger range
cat(perc,perc_am,perc_iucn)

@@

>>>>

spp_list = spp_list%>%
              mutate(total_area=NA,
                     am_area = NA,
                     iucn_area=NA,
                     perc = NA,
                     am_perc=NA,
                     iucn_perc=NA,
                     sm_perc = NA, #filled with % of smaller range that is within the larger range
                     sm_range = NA) #what range is smaller

@@

  #get just the aquamaps cells
#   am_spp_map = spp_map%>%
#                   filter(!is.na(am_prob))%>%
#                 dplyr::select(loiczid,am_sid,am_prob)%>% #add in am_prob when we want to look at probability
#                   unique()

#   if(TRUE %in% duplicated(am_spp_map$loiczid)){
#                   am_spp_map = am_spp_map%>%
#                                 dplyr::select(loiczid,am_sid)%>%
#                                   unique()}
#   
  #rasterize aquamaps map
#   r_am_spp  <-  subs(loiczid_raster, 
#                      am_spp_map[ , c('loiczid', 'am_sid')], 
#                      by = 'loiczid', 
#                      which = 'am_sid', 
#                      subsWithNA = TRUE)
#   
#   spp_list[i,]$am_area = area(r_am_spp,na.rm=T)%>%
#               cellStats(.,stat='sum')
#

  #get the iucn cells
#   iucn_map <- spp_map %>%
#     mutate(iucn_pres = iucn_area > 0)%>%
#       dplyr::select(loiczid,iucn_area,am_sid,am_prob,iucn_pres)%>%
#         unique()

@@

#    if(TRUE %in% duplicated(iucn_map$loiczid))next()
# #   
#  rasterize IUCN map
#   r_iucn_spp <- subs(loiczid_raster, 
#                      iucn_map[ , c('loiczid', 'iucn_pres')], 
#                     by = 'loiczid', 
#                      which = 'iucn_pres', 
#                      subsWithNA = TRUE)
# #   
#   spp_list[i,]$iucn_area = area(r_iucn_spp,na.rm=T)%>%
#                 cellStats(.,stat='sum')

  # Get total area (IUCN and AM combined)
  allCells <- spp_map%>%
                dplyr::select(loiczid)%>%
                  mutate(value=1)%>%
                  unique()

  fullRange <- subs(loiczid_raster, 
                     allCells[ , c('loiczid', 'value')], 
                    by = 'loiczid', 
                     which = 'value', 
                     subsWithNA = TRUE)

  spp_list[i,]$total_area = area(fullRange,na.rm=T)%>%
                              cellStats(.,stat='sum')

@@

#   #iucn map cells
#   cells_iucn<-iucn_map%>%
#               filter(iucn_pres==TRUE)%>%
#               .$loiczid
#   cells_am <- am_spp_map%>%
#                 filter(!is.na(am_sid))%>%
#                 .$loiczid
#   
#   cells_total = unique(spp_map$loiczid)
#   cells_overlap = intersect(cells_iucn,cells_am)
#   
#    r_iucn_spp <- subs(loiczid_raster, 
#                       iucn_map[ , c('loiczid', 'iucn_pres')], 
#                      by = 'loiczid', 
#                       which = 'iucn_pres', 
#                       subsWithNA = TRUE)
#   
  #percent overlap
#   perc = (length(cells_overlap)/length(cells_total))*100
#   spp_list[i,]$perc = perc
#   
#   am_only = setdiff(cells_am,cells_iucn)
#   iucn_only = setdiff(cells_iucn,cells_am)
#   
#   perc_am = (length(am_only)/length(cells_total))*100
#   perc_iucn = (length(iucn_only)/length(cells_total))*100
#   
#   spp_list[i,]$am_perc = perc_am
#   spp_list[i,]$iucn_perc = perc_iucn
# 
#   spp_list[i,]$sm_range = ifelse(length(cells_am)<length(cells_iucn),'AM','IUCN') #what dataset has the smaller range
#   spp_list[i,]$sm_perc = length(cells_overlap)/min(length(cells_am),length(cells_iucn))*100 #this is the total percent of smaller range encompassed in the larger range
# cat(perc,perc_am,perc_iucn)
#

@@

reg <- ggplot(spp_list, aes(sm_perc,total_area,color=sm_range)) + geom_point() + geom_abline() + labs(y='Total Area', x='Percent overlap smaller range in larger', title='Comparing differences in area across species groups', fill='IUCN Category')

print(reg)

model = lm(spp_list$total_area~spp_list$sm_perc)
summary(model)
